---
title: "Membaca File xls dan xlsx"
output: html_notebook
---

### Membaca file xls

Gunakan library readxl

```{r}
library(readxl)
myRsample <- read_excel("myRsample.xls")

```


### Membaca file xlsx

Gunakan library xlsx

```{r}
library("xlsx")

read.xlsx("myRsample.xls", sheetIndex, header=TRUE, colClasses=NA)
read.xlsx2("myRsample.xls", sheetIndex, header=TRUE, colClasses="character")

```

### Menulis file xlsx

Gunakan library xlsx

```{r}
write.xlsx(x, file, sheetName="Sheet1", col.names=TRUE, row.names=TRUE, append=FALSE)
write.xlsx2(x, file, sheetName="Sheet1", col.names=TRUE, row.names=TRUE, append=FALSE)
```

x : data.frame
file : path output file
sheetName : a character string to use for the sheet name.
col.names, row.names : a logical value specifying whether the column names/row names of x are to be written to the file
append : a logical value indicating if x should be appended to an existing file.


http://www.sthda.com/english/wiki/r-xlsx-package-a-quick-start-guide-to-manipulate-excel-files-in-r